---
title: "caracas: Computer Algebra via SymPy"
output: html_notebook
---
https://mikl.dk/post/2020-caracas_v101/

It is with great pleasure that we can announce the release of caracas version 1.0.1 to CRAN (https://cran.r-project.org/package=caracas).

The package enables user to make computer algebra from R using the Python library SymPy.

You can now install the caracas package as follows:

install.packages("caracas")
And then load it by:

library(caracas)
The source code and the development version is available at https://github.com/r-cas/caracas/. Online documentation (of the development version) can be found at https://r-cas.github.io/caracas/.

Below we will show a few examples, but also notice the vignettes available at https://cran.r-project.org/package=caracas.


```{r}
library(caracas)
library(reticulate)
```

Additional steps needed not in article. This is needed to get reticulate to work correctly and to get the packages installed with conda. Link to the article explaining this: https://rstudio.github.io/reticulate/articles/python_packages.html

```{r}
conda_create("r-reticulate")
conda_install("r-reticulate","sympy")

```
Not in article - this puts sympy into the r-space
```{r}
#reticulate::use_codaenv("r-reticulate")
sympy<-import("sympy")
```

And now the article continues: 
Quick start:
```{r}
x <- symbol('x')
eq <- 2*x^2 - x
eq

cat("As character string:\n")
as.character(eq)

cat("\nAs R expression:\n")
as_r(eq)

cat("\nAnd as a tex object:\n")
tex(eq)


```
$2 x^{2} - x$

```{r}
solve_sys(eq,x)


der(eq,x)

subs(eq,x, "y")
```

#Linear Algebra#

```{r}

A <- matrix(c("x", 2, 0, "2*x"), 2, 2)
B <- as_symbol(A)
B


```

```{r}
determinant(B)

Binv<- inv(B)
Binv

tex(Binv)
```
$\\left[\\begin{matrix}\\frac{1}{x} & 0\\\\- \\frac{1}{x^{2}} & \\frac{1}{2 x}\\end{matrix}\\right]$

$[\begin{matrix}\frac{1}{x} & 0\\- \frac{1}{x^{2}} & \frac{1}{2 x}\end{matrix}]$

```{r}
eigen_val(Binv)

```

#Maximising the multinomial likelihood#

```{r}
p <- as_symbol(paste0("p", 1:3))
y <- as_symbol(paste0("y", 1:3))
a <- as_symbol("a")
l <- sum(y*log(p))
l

```

```{r}
L <- -l + a*(sum(p) - 1)
L
#
g <- der(L, c(p, a))
g
#
sol <- solve_sys(g, c(p, a))
sol
#
sol[[1L]]$p1

#
tex(sol[[1L]]$p1)



```
$\frac{y_{1}}{y_{1} + y_{2} + y_{3}}$

```{r}
l
l_e <-as_r(l)

eval(l_e, list(p1 = 0.2, p2 = 0.3, p3 = 0.5, 
               y1 = 18, y2 = 31, y3 = 51))
```

More information and examples
Please refer to one or more of the following:

https://cran.r-project.org/package=caracas
https://github.com/r-cas/caracas/
https://r-cas.github.io/caracas/
